{Contextual Learning with Integrated Least Effort Knowledge Networks}

### 2016-10-20 proposal.tex
proposal final draft

\renewcommand{\baselinestretch}{1.2}

\title{{\Large Contextual Learning with Integrated Least Effort Knowledge Networks}}
\author{Lucas E. Morales \texttt{\{lucasem\}}\\
Supervised by: Joshua Tenenbaum \texttt{\{jbt\}}}
\date{}

@@

@@

This project introduces the SKN framework ({\bf S}cale-free {\bf K}nowledge
{\bf N}etwork), capable of contextualizing knowledge in an abstract manner.
Knowledge can be represented as distinct atoms of information and their
relations in the structure of a connected network. The network is
constructed by preferential ``least effort'' attachment
\cite{barabasi99}\cite{cancho03}, where the learning of a new atom of
knowledge joins the network with relations to the contextual knowledge from
which it was learned based on memory access. This network is scale-free ---
it conforms to a mathematical pattern similar to that of Zipf's Law
\cite{zipf49}, where the degree of connectivity for nodes in the network
follow a class of power law distributions.

@@

\begin{figure}[ht]
\centering
\includegraphics[scale=.5]{sf_add.eps}
\caption{The first pane shows a knowledge network, and a shaded region
  representing the active context. The second pane shows the memory accesses
  to particular items of knowledge leading up to the addition of a new item
  of knowledge. The third pane shows the new knowledge network, with the
  additional item of knowledge.}
\label{fig:sf_add}
\end{figure}

@@

@@

To experiment with SKN, consider the class of problems for which this system
would excel: complex domains where modular knowledge improves the efficiency
of completing relevant tasks. We augment a system for learning via iterative
modular concept discovery --- the EC algorithm \cite{dechter13} --- to
interface with SKN rather than use a flat set of knowledge items. The
augmented EC algorithm will be tested in the domains of letter analogies
\cite{hofstadter94}, L-systems \cite{rozenberg80}, and Boolean circuits
illustrated in Figure \ref{fig:exps}, and compared to the original EC algorithm in the as
well as a baseline of search without knowledge assistance. These problem
domains were chosen because they each have potential for common underlying
structure between tasks. Results will be compared on the number of tasks
solved given number of iterations for the algorithm.

\begin{figure}[h]
\centering
\includegraphics[scale=.5]{exps.eps}
\caption{Domains for test experiments.}
\label{fig:exps}
\end{figure}

@@

\begin{itemize}
  \item September-October --- preliminary formalization, decide on
    experiments.
  \item November-January --- preliminary implementation, completed
    formalization.
  \item February --- finalize implementation.
  \item March-April --- complete paper writeup.
\end{itemize}

>>>>

\renewcommand{\baselinestretch}{1.3}

\title{{Contextual Learning with Integrated Least Effort Knowledge Networks}}
\author{Lucas E. Morales \texttt{\{lucasem\}}\\
Principal Invesigator: Joshua Tenenbaum \texttt{\{jbt\}}}
\date{}

@@

\begin{abstract}
  In the field of artificial intelligence, most learning mechanisms are
  either inherently specialized or have a flat knowledge base that makes
  them only useful when confined to a single domain, which is flawed as a
  model of cognition because it reduces capability of both handling very
  complex problems and specializing in multiple domains. This project is to
  design a contextual knowledge system called Scale-free Knowledge Network
  (SKN). We augment learning mechanisms to utilize this knowledge system in
  a manner that reduces the intractability of hard problems by creating
  implicit relationships between items of knowledge.
\end{abstract}

@@

We present the SKN framework ({\bf S}cale-free {\bf K}nowledge {\bf
N}etwork), capable of contextualizing knowledge in an abstract manner.
Knowledge can be represented as distinct atoms of information and their
relations in the structure of a connected network. The network is
constructed by preferential ``least effort'' attachment
\cite{barabasi99}\cite{cancho03}, where the learning of a new atom of
knowledge joins the network with relations to the contextual knowledge from
which it was learned based on memory access. This network is scale-free ---
it conforms to a mathematical pattern similar to that of Zipf's Law
\cite{zipf49}, where the degree of connectivity for nodes in the network
follow a class of power law distributions.

@@

\begin{figure}[h]
\centering
\includegraphics[scale=.5]{sf_add.eps}
\caption{The first pane shows a knowledge network, and a shaded region
  representing the active context. The second pane shows the memory accesses
  to particular items of knowledge leading up to the addition of a new item
  of knowledge. The third pane shows the new knowledge network, with the
  additional item of knowledge.}
\label{fig:sf_add}
\end{figure}

@@

\begin{figure}[h]
\centering
\includegraphics[scale=.5]{exps.eps}
\caption{Domains for test experiments. Items in each of these three domains
  have substructures of modular concepts which are useful across many tasks,
  making them fitting domains to utilize contextual knowledge.}
\label{fig:exps}
\end{figure}

@@

To experiment with SKN, consider the class of problems for which this system
would excel: complex domains where modular knowledge improves the efficiency
of completing relevant tasks. We augment a system for learning via iterative
modular concept discovery --- the EC algorithm \cite{dechter13} --- to
interface with SKN rather than use the traditional flat set of knowledge
items. The EC algorithm performs search on a stochastic grammar that is
expressive in a particular problem domain, and modifies the grammar's
generative distribution to include learned modular concepts which are useful
in representing tasks. The augmentation to utilize SKN will treat the
generative distribution as a mixture model that uses concepts which are
stored in the context. The augmented EC algorithm will be tested in the
domains of letter analogies \cite{hofstadter94}, L-systems
\cite{rozenberg80}, and Boolean circuits illustrated in Figure
\ref{fig:exps}, and compared to the original EC algorithm as well as a
baseline of search without knowledge assistance. These problem domains were
chosen because they each have potential for common underlying structure
between tasks. Results will be compared on the number of tasks solved given
number of iterations for the algorithm.

@@

\begin{itemize}
  \item September-October --- preliminary formalization, decide on
    experiments.
  \item November-December --- preliminary implementation, completed
    formalization.
  \item January-February --- finalize implementation, complete
    experimentation.
  \item March-April --- prepare writeup for submission to the AGI
    conference, as well as a workshop for the UAI or IJCAI conferences.
\end{itemize}
